---
title: "text_outpus"
output: word_document
date: "2025-09-04"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load_packages, echo = FALSE}
# function used to load packages current session while suppressing warnings and start up messages
load_package <- function(pckg) {
  suppressWarnings(suppressPackageStartupMessages(library(pckg, character.only = T, warn.conflicts = F, verbose = F, quietly = T)))
}

# Load packages, suppressing warnings so easier when batch run
pckgs <- c("tidyverse", "jsonlite")
sapply(pckgs, function(pckg) { load_package(pckg) } )
options(dplyr.summarise.inform = F)
```

```{r functions, echo=FALSE}
current_script <- "markdown"
# helper functions and global variables
configs <- fromJSON("../common_scripts/00_configs.json")

## source("../common_scripts/00_global_vars.R")
```

```{r global_var, echo = FALSE}
# Locations where to read data from 
results_generated_dt_main <- configs$main_configs$results_generated_dt
results_loc_main          <- configs$main_configs$results_loc
results_loc_main_suppress <- str_glue("{results_loc_main}suppressed/")

results_generated_dt_sa_pc <- configs$sa_configs$results_generated_dt
results_loc_sa_pc          <- paste0(configs$sa_configs$results_loc, "prostate_cancer/")
results_loc_sa_pc_suppress <- str_glue("{results_loc_sa_pc}suppressed/")

results_generated_dt_sa_pc <- configs$sa_configs$results_generated_dt
results_loc_sa_rtd <- paste0(configs$sa_configs$results_loc, "rtd/")
results_loc_sa_rtd_suppress <- str_glue("{results_loc_sa_rtd}/suppressed/")

symptom_groups   <- read.delim(paste0(here::here(), "/lookup_files/symptom_groups.txt"), sep = " ")

save_dir    <- paste0(here::here(), "/tables_figs/manuscript/")
if(!dir.exists(save_dir)) dir.create(save_dir)

save_dir_suppress  <- paste0(save_dir, "suppressed/")
if(!dir.exists(save_dir_suppress)) dir.create(save_dir_suppress)
```

# Main analysis

## Stage distribution by symptom observations

```{r stage_sx, echo = FALSE}
sx_order                 <- readRDS(paste0(results_loc_main, "order_stage_by_sex_symptom_", results_generated_dt_main, ".rds"))
stage_bin_by_sex_symptom <- read.csv(paste0(results_loc_main, "stage_bin_by_sex_symptom_", results_generated_dt_main, ".csv"))
stage_bin_by_sex_symptom <- stage_bin_by_sex_symptom |> left_join(symptom_groups)

stage_bin_by_sex_symptom <- stage_bin_by_sex_symptom |>
  filter(N_tumour >= 100) |>
  pivot_wider(id_cols = c(sex_lb, symptom_group, symptom),
              names_from = stage_bin,
              values_from = c(prop100, prop100_lb, prop100_ub))

stage_bin_by_sex_symptom <- 
  stage_bin_by_sex_symptom |> 
  group_by(sex_lb) |> 
  arrange(desc(`prop100_1-3`)) |> 
  ungroup()

stage_bin_by_sex_symptom |> 
  filter(symptom_group %in% c("Skin Lesion", "Male specific", "Female specific", "Breast Symptoms")) |>
  select(-ends_with("4"))

stage_bin_by_sex_symptom |> 
  filter(grepl(".*pain.*", symptom)) |>
  select(-ends_with("4")) |>
  arrange(desc(`prop100_1-3`))
```
## Suppressed: Stage distribution by symptom observations

```{r stage_sx_suppress, echo = FALSE}
sx_order                 <- readRDS(paste0(results_loc_main, "order_stage_by_sex_symptom_", results_generated_dt_main, ".rds"))
stage_bin_by_sex_symptom_suppress <- read.csv(paste0(results_loc_main_suppress, "stage_bin_by_sex_symptom_", results_generated_dt_main, ".csv"))
stage_bin_by_sex_symptom_suppress <- stage_bin_by_sex_symptom_suppress |> left_join(symptom_groups)

# based on actual number of tumours
stage_bin_by_sex_symptom_suppress <-
  stage_bin_by_sex_symptom_suppress |>
  filter(N_tumour >= 100)

# clean dataframe 
stage_bin_by_sex_symptom_suppress <-
  stage_bin_by_sex_symptom_suppress |>
  select(
    sex_lb, symptom_group, symptom, stage_bin,
    prop100    = prop100_suppressed,
    prop100_lb = prop100_lb_min,
    prop100_ub = prop100_ub_max
  )

stage_bin_by_sex_symptom_suppress <- 
  stage_bin_by_sex_symptom_suppress |>
  pivot_wider(id_cols = c(sex_lb, symptom_group, symptom),
              names_from = stage_bin,
              values_from = c(prop100, prop100_lb, prop100_ub))

stage_bin_by_sex_symptom_suppress <- 
  stage_bin_by_sex_symptom_suppress |> 
  group_by(sex_lb) |> 
  arrange(desc(`prop100_1-3`)) |> 
  ungroup()

stage_bin_by_sex_symptom_suppress |> 
  filter(symptom_group %in% c("Skin Lesion", "Male specific", "Female specific", "Breast Symptoms")) |>
  select(-ends_with("4"))

stage_bin_by_sex_symptom_suppress |> 
  filter(grepl(".*pain.*", symptom)) |>
  select(-ends_with("4")) |>
  arrange(desc(`prop100_1-3`))
```

## stage distribution by symptom, stratified by cancer site

```{r stage_sx_site, echo = FALSE}
stage_bin_by_symptom_site_exemplar <- read.csv(str_glue("{results_loc_main}stage_bin_by_symptom_site_exemplar_{results_generated_dt_main}.csv"))

stage_bin_by_symptom_site_exemplar |> distinct(cancer_site_desc)

stage_bin_by_symptom_site_exemplar |> 
  filter(symptom  == "Weight loss") |>
  filter(stage_bin == "1-3") |>
  arrange(desc(prop100))
```
## Suppressed: Stage distribution by symptom, stratified by cancer site

```{r stage_sx_site_suppress, echo = FALSE}
stage_bin_by_symptom_site_exemplar_suppress <- read.csv(str_glue("{results_loc_main_suppress}stage_bin_by_symptom_site_exemplar_{results_generated_dt_main}.csv"))

stage_bin_by_symptom_site_exemplar_suppress |> distinct(cancer_site_desc)

stage_bin_by_symptom_site_exemplar_suppress <-
  stage_bin_by_symptom_site_exemplar_suppress |>
  select(symptom, cancer_site_desc, stage_bin,
         n_tumours  = suppressed_n_tidy,
         N_tumours  = suppressed_N,
         prop100    = prop100_suppressed,
         prop100_lb = prop100_lb_min,
         prop100_ub = prop100_ub_max)

stage_bin_by_symptom_site_exemplar_suppress |> 
  filter(symptom  == "Weight loss") |>
  filter(stage_bin == "1-3") |>
  arrange(desc(prop100))
```

## Stage distribution by symptom, stratified by blood test use

```{r stage_sx_bt, echo = FALSE}
stage_bin_by_sex_symptom_bt  <- read.csv(paste0(results_loc_main, "stage_bin_by_sex_symptom_bt_", results_generated_dt_main, ".csv")) 
t_test                       <- read.csv(paste0(results_loc_main, "t_test_bt_",  results_generated_dt_main, ".csv")) 

t_test |> group_by(sex_lb, signif) |> summarise(n = n())

stage_bin_by_sex_symptom_bt |>
  inner_join(t_test |> filter(signif), by = c("sex_lb", "symptom")) |>
  filter(stage_bin == 4) |>
  pivot_wider(id_cols     = c(sex_lb, symptom),
              names_from  = blood_test,
              values_from = starts_with("prop100"), 
              names_vary  = "slowest")  |>
  arrange(sex_lb, desc(prop100_0)) |>
  mutate(stage4_bt_higher_no_bt = prop100_1 > prop100_0) 

stage_bin_by_sex_symptom_bt |>
  inner_join(t_test |> filter(signif), by = c("sex_lb", "symptom")) |>
  filter(stage_bin == 4) |>
  pivot_wider(id_cols     = c(sex_lb, symptom),
              names_from  = blood_test,
              values_from = starts_with("prop100"), 
              names_vary  = "slowest")  |>
  mutate(abs_diff = abs(prop100_0 - prop100_1)) |>
  arrange(sex_lb, abs_diff)
```
## Suppressed: Stage distribution by symptom, stratified by blood test use

```{r stage_sx_bt_suppress, echo = FALSE}
stage_bin_by_sex_symptom_bt_suppress  <- read.csv(str_glue("{results_loc_main_suppress}stage_bin_by_sex_symptom_bt_{results_generated_dt_main}.csv")) 
t_test                       <- read.csv(paste0(results_loc_main, "t_test_bt_",  results_generated_dt_main, ".csv")) 

t_test |> group_by(sex_lb, signif) |> summarise(n = n())

stage_bin_by_sex_symptom_bt_suppress <-
  stage_bin_by_sex_symptom_bt_suppress |>
  select(sex_lb, symptom, blood_test, stage_bin, prop100 = prop100_suppressed, prop100_lb = prop100_lb_min, prop100_ub, prop100_ub_max)

stage_bin_by_sex_symptom_bt_suppress |>
  inner_join(t_test |> filter(signif), by = c("sex_lb", "symptom")) |>
  filter(stage_bin == 4) |>
  pivot_wider(id_cols     = c(sex_lb, symptom),
              names_from  = blood_test,
              values_from = starts_with("prop100"), 
              names_vary  = "slowest")  |>
  arrange(sex_lb, desc(prop100_0)) |>
  mutate(stage4_bt_higher_no_bt = prop100_1 > prop100_0) 

stage_bin_by_sex_symptom_bt_suppress |>
  inner_join(t_test |> filter(signif), by = c("sex_lb", "symptom")) |>
  filter(stage_bin == 4) |>
  pivot_wider(id_cols     = c(sex_lb, symptom),
              names_from  = blood_test,
              values_from = starts_with("prop100"), 
              names_vary  = "slowest")  |>
  mutate(abs_diff = abs(prop100_0 - prop100_1)) |>
  arrange(sex_lb, abs_diff)
```

## Stage distribution by symptom combinations, stratified by sex and cancer site

```{r stage_sx_comb, echo = FALSE}
model_tidy <- read.csv(str_glue("{results_loc_main}model_new_symptom_combinations_{results_generated_dt_main}.csv"))
head(model_tidy)

model_tidy |> filter(grepl(".*\\:.*", term)) |> filter(p < 0.05) |> select(sex, cancer_site_desc, term, p) |>
  mutate(p = round(p, 5))
```

# Sensitivity analysis - exluding prostate cancer

## Stage distribution by symptom, stratified by blood test use (excluding prostate cancer)

```{r sa_stage_sx_bt, echo = FALSE}
stage_bin_by_sex_symptom_bt_sa  <- read.csv(paste0(results_loc_sa_pc, "stage_bin_by_sex_symptom_bt_", results_generated_dt_sa_pc, ".csv")) 
t_test                          <- read.csv(paste0(results_loc_main, "t_test_bt_",  results_generated_dt_main, ".csv")) 

t_test |> group_by(sex_lb, signif) |> summarise(n = n())

stage_bin_by_sex_symptom_bt_sa |>
  inner_join(t_test |> filter(signif), by = c("sex_lb", "symptom")) |>
  filter(stage_bin == 4) |>
  pivot_wider(id_cols     = c(sex_lb, symptom),
              names_from  = blood_test,
              values_from = starts_with("prop100"), 
              names_vary  = "slowest")  |>
  arrange(sex_lb, desc(prop100_0)) |>
  mutate(stage4_bt_higher_no_bt = prop100_1 > prop100_0) 

stage_bin_by_sex_symptom_bt_sa |>
  inner_join(t_test |> filter(signif), by = c("sex_lb", "symptom")) |>
  filter(stage_bin == 4) |>
  pivot_wider(id_cols     = c(sex_lb, symptom),
              names_from  = blood_test,
              values_from = starts_with("prop100"), 
              names_vary  = "slowest")  |>
  mutate(abs_diff = abs(prop100_0 - prop100_1)) |>
  arrange(sex_lb, abs_diff)
```
## Suppressed: Stage distribution by symptom, stratified by blood test use (excluding prostate cancer)

```{r sa_stage_sx_bt_suppress, echo = FALSE}
stage_bin_by_sex_symptom_bt_sa_suppress  <- read.csv(str_glue("{results_loc_sa_pc_suppress}stage_bin_by_sex_symptom_bt_{results_generated_dt_sa_pc}.csv")) 
t_test                          <- read.csv(paste0(results_loc_main, "t_test_bt_",  results_generated_dt_main, ".csv")) 

t_test |> group_by(sex_lb, signif) |> summarise(n = n())

stage_bin_by_sex_symptom_bt_sa_suppress <-
  stage_bin_by_sex_symptom_bt_sa_suppress |>
  select(sex_lb, symptom, blood_test, stage_bin, prop100 = prop100_suppressed,
         prop100_lb = prop100_lb_min,
         prop100_ub = prop100_ub_max)

stage_bin_by_sex_symptom_bt_sa_suppress |>
  inner_join(t_test |> filter(signif), by = c("sex_lb", "symptom")) |>
  filter(stage_bin == 4) |>
  pivot_wider(id_cols     = c(sex_lb, symptom),
              names_from  = blood_test,
              values_from = starts_with("prop100"), 
              names_vary  = "slowest")  |>
  arrange(sex_lb, desc(prop100_0)) |>
  mutate(stage4_bt_higher_no_bt = prop100_1 > prop100_0) 

stage_bin_by_sex_symptom_bt_sa_suppress |>
  inner_join(t_test |> filter(signif), by = c("sex_lb", "symptom")) |>
  filter(stage_bin == 4) |>
  pivot_wider(id_cols     = c(sex_lb, symptom),
              names_from  = blood_test,
              values_from = starts_with("prop100"), 
              names_vary  = "slowest")  |>
  mutate(abs_diff = abs(prop100_0 - prop100_1)) |>
  arrange(sex_lb, abs_diff)
```
## Suppressed: Route to diagnosis (symptomatic vs non-symptomatic)
```{r sa_rtd}
N_tumour_by_sex_symptom_rtd_suppress <- read.csv(
  str_glue("{results_loc_sa_rtd_suppress}n_tumour_by_sex_symptom_rtd_{results_generated_dt_sa_pc}.csv"))
N_tumour_by_sex_symptom_rtd_suppress |>
  filter(N_tumour >= 100) |>
  mutate(unknown_sx = symptom %in% c("N/A", "N/K")) |>
  group_by(unknown_sx, typereferral_recode) |>
  summarise(n_tumour = sum(n_tumour), .groups = "drop_last") |>
  mutate(N_tumour = sum(n_tumour),
         prop100 = 100*n_tumour/N_tumour) |>
  filter(typereferral_recode == "Unknown")
```
